Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-11-12 main.tex
Addressed the comment about direction of perturbation mattering.

We record the order number of each perturbation \(j\), i.e. first, second,
third, \ldots, to measure how many perturbations a rider is subjected to before
the current perturbation. At the initiation of each perturbation we log the
instantaneous steer and roll angles, \(\delta_0\) and \(\phi_0\), to
characterize the configuration of the bicycle when perturbed. The gain factor
setting on the balance assist controller indicates if the assistance is off
\(\kappa=0\) or on at two different levels: low \(\kappa=3.9\) or high
\(\kappa=5.2\). As mentioned earlier, a recovery from the perturbation is
successful if the person neither places their foot down onto the treadmill
surface nor their wheel of the bicycle exits the width of the treadmill belt.
We record this as a binary variable \(f\) for ``fall''.  All measured variables
are reported in Table~\ref{tab:raw}.
%
\begin{table}
  \centering
  \caption{Raw measurements taken during each trial. ``Fall Outcome'' and
    ``Perturbation Order Number'' are recorded per perturbation instance. The
    remaining measures are time varying during each trial.}
  \begin{tabular}{llll}
    \toprule
    Measure & Variable & Units & Sensor \\
    \midrule
    Balance Assist Gain Factor & \(\kappa\) & \si{\newton\second\squared} & NA \\
    Bicycle Speed & \(v\) & \si{\meter\per\second} &  DF30 wheel encoder \\
    Fall Outcome & \(f\) & boolean & NA \\
    Force \(l\)eft/\(r\)ight,\(f\)ront/\(r\)ear &
    \(F_{lf},F_{rf},F_{lr},F_{rr}\) & \si{\newton} & inline load cell\\
    Perturbation Order Number & \(j\) & integer & NA\\
    Roll Angle & \(\phi\) & \si{\degree} & BN0086 Kalman estimate \\
    Roll Angular Rate & \(\dot{\phi}\) & \si{\degree\per\second} & BN0086 rate gyroscope \\
    Steer Angle & \(\delta\) & \si{\degree} & steer tube encoder \\
    \bottomrule
  \end{tabular}
  \label{tab:raw}
\end{table}

>>>>

We record the order number of each perturbation \(j\), i.e. first, second,
third, \ldots, to measure how many perturbations a rider is subjected to before
the current perturbation. At the initiation of each perturbation we log the
instantaneous steer and roll angles, \(\delta_0\) and \(\phi_0\), to
characterize the configuration of the bicycle when perturbed. The sense of
these variables relative to the sense of the perturbation matters. For example,
if there is already a large steer angle in the same direction as the
perturbation, this will have a different effect on the probability that a fall
will occur than if the steer angle is in the opposite direction of the
perturbation. To account for this, the roll and steer variables at the instance
of perturbation are normalized for perturbation direction. The gain factor
setting on the balance assist controller indicates if the assistance is off
\(\kappa=0\) or on at two different levels: low \(\kappa=3.9\) or high
\(\kappa=5.2\). As mentioned earlier, a recovery from the perturbation is
successful if the person neither places their foot down onto the treadmill
surface nor their wheel of the bicycle exits the width of the treadmill belt.
We record this as a binary variable \(f\) for ``fall''.  All measured variables
are reported in Table~\ref{tab:raw}.
%
\begin{table}
  \centering
  \caption{Raw measurements taken during each trial. ``Fall Outcome'' and
    ``Perturbation Order Number'' are recorded per perturbation instance. The
    remaining measures are time varying during each trial.}
  \begin{tabular}{llll}
    \toprule
    Measure & Variable & Units & Sensor \\
    \midrule
    Balance Assist Gain Factor & \(\kappa\) & \si{\newton\second\squared} & NA \\
    Bicycle Speed & \(v\) & \si{\meter\per\second} &  DF30 wheel encoder \\
    Fall Outcome & \(f\) & boolean & NA \\
    Force \(l\)eft/\(r\)ight,\(f\)ront/\(r\)ear &
    \(F_{lf},F_{rf},F_{lr},F_{rr}\) & \si{\newton} & inline load cell\\
    Perturbation Order Number & \(j\) & integer & NA\\
    Roll Angle & \(\phi\) & \si{\degree} & BN0086 Kalman estimate \\
    Roll Angular Rate & \(\dot{\phi}\) & \si{\degree\per\second} & BN0086 rate gyroscope \\
    Steer Angle & \(\delta\) & \si{\degree} & steer tube encoder \\
    \bottomrule
  \end{tabular}
  \label{tab:raw}
\end{table}
